\Large Analyzing the Supreme Court Citation Network

### 2017-12-28 Tex/paper.tex
revising model definition

\textbf{Sociality} Sociality is the tendency for ties sent to beget more ties sent [CITE]. Structurally, this means that there are few nodes that send a moderate number of ties---the number of ties sent is either small or fairly large. In terms of Supreme Court opinions,  for each opinion discussed, that discussion is likely to raise other tangential issues on which the justices will want to draw upon past opinions.  Furthermore, for each opinion that applies to and is cited by the current opinion, there is often a case/opinion that needs to be discussed in terms of why it does not apply. Justices often clarify not only those legal principles that apply, but often those that do not. When multiple lines of legal reasoning are drawn upon to justify a decision, often more are discussed in terms of their lack of applicability.

@@

A citation is a directed link between two documents that indicates the citing document attributes the cited to be in some manner relevant to it. Collective patterns of citations within a certain domain make up citation networks, which are most commonly found in three social systems, scientific publishing, patents, and the judicial system. When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc, usually unweighted, is the existence of one or more cites from the sending to the receiving document. Given the permanent nature of documents in citations networks, these networks are in large acyclic \citep{leicht2007large,karrer2009random}. Relatedly, citation networks grow as new documents and citations enter the network, but established arcs persist in all but exceptional cases.

@@

 \section{The Exponential Random Configuration Model}
Let $c(t)\in \{0,1\}^N$ be a vector indicating which Supreme Court case has been cited at time $t$, where $c_i(t)=1, i \in \{1, \dots , N\}$ indicates that the $i$th case has been cited at time $t$ and $c_i(t)=0$ indicates that the $i$th case has not been cited at time $t$. Furthermore, let
$$\mathcal{C}_t(N)=\{c(t)\in \{0,1\}^N: c_i(t)\in \{0,1\} \}$$ 
be the set of all possible citation combinations at time $t$. Note that the cardinality of $\mathcal{C}_t(N)$ increases exponentially for every newly added case, which results in $2^N$ elements.\\
The probability function of the ERCM is defined as
\begin{equation}
P_{\theta}(c(t)~|~c(t-))=\cfrac{\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}{\sum_{c(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}
\label{ercm}
\end{equation}
where $c(t-)\in \{0,1\}^{N \times (t-1)}$ is a matrix that indicates which cases have been citing in each other before time $t$, $\theta \in \mathbb{R}^q$ is a q-dimensional vector of parameters,  $h: \mathcal{C}_t(N) \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{c(t)^*\in \mathcal{C}}\exp(\theta^T\cdot h(c(t)|c(t-)))$ is a normalization constant that ensures that (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.\\[0.3cm]

@@

Our case for studying legal citaions at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence. To quantify this contribution, we use out-of-sample prediction [CITES]. Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model. Unlike in-sample measures of model fit, out-of-sample methods are highly robust in avoiding overfitting, and work when we cannot accurately calculate the value of the likelihood function, as in the current case.

>>>>

\textbf{Sociality} Sociality is the tendency for ties sent to beget more ties sent [CITE]. Structurally, this means that there are few nodes that send a moderate number of ties---the number of ties sent is either small or fairly large. In terms of Supreme Court opinions,  for each opinion discussed, that discussion is likely to raise other tangential issues on which the justices will want to draw upon past opinions.  Furthermore, for each opinion that applies to and is cited by the current opinion, there is often a case/opinion that needs to be discussed in terms of why it does not apply. Justices often clarify not only those legal principles that apply, but often those that do not. The passages below represent examples of this process of citation proliferation, or sociality.

@@

Researchers from several fields have used network analysis to analyze citations---legal, patent, and scientific. Before describing the network model for Supreme Court citations that we develop, we review the methods researchers have previously used to study citations.. A citation is a directed link between two documents that indicates the citing document attributes the cited to be in some manner relevant to it. Collective patterns of citations within a certain domain make up citation networks, which are most commonly found in three social systems, scientific publishing, patents, and the judicial system. When raw citation data is formally represented as a network, the nodes are usually the documents themselves, and each directed arc, usually unweighted, is the existence of one or more cites from the sending to the receiving document. Given the permanent nature of documents in citations networks, these networks are in large acyclic \citep{leicht2007large,karrer2009random}. Relatedly, citation networks grow as new documents and citations enter the network, but established arcs persist in all but exceptional cases.

@@

 \section{The Citation Temporal Exponential Random Graph Model}
We build upon existing methods for network analysis to define a model that fits the constraints to which Supreme Court citation networks are subject. The Supreme Court citation network is nearly a directed acyclic graph---a graph in which there can be no loops/cycles (e.g., ties cannot be reciprocated, which would constitute a two-node loop). If two cases are decided in different terms, the later case can cite the earlier case, but not the reverse. However, as noted above, if two opinions are written in the same term, they can cite each other reciprocally. We implement a version of the temporal exponential random graph model (TERGM)---a model that is commonly used for longitudinal network data [CITES]---that is subject to the constraints specific to the Supreme Court citation network. The TERGM is a model that can simultaneously incorporate structural dependencies such as transitivity and reciprocity and exogenous covariates  (e.g. the issue area of the citing case, the ideological distance between the opinion authors of the two cases in the dyad) to explain tie formation. If the ties do not depend upon each other, and the coefficients associated with the structural dependencies are either assumed or estimated to be zero, the TERGM reduces to a logistic regression in which the observation is a directed dyad [CITE OLD PA].  In this section we describe the citation TERGM (c-TERGM), and explain estimation methods.

Let $c_ij \in \{0,1\}$ be a binary indicator of whether case $i$ cites $j$. Furthermore, let
$$\mathcal{C}_t(N)=\{c(t)\in \{0,1\}^N: c_i(t)\in \{0,1\} \}$$ 
be the set of all possible citation combinations at time $t$. Note that the cardinality of $\mathcal{C}_t(N)$ increases exponentially for every newly added case, which results in $2^N$ elements.\\
The probability function of the ERCM is defined as
\begin{equation}
P_{\theta}(c(t)~|~c(t-))=\cfrac{\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}{\sum_{c(t)^*\in \mathcal{C}}\exp\big(\theta^T\cdot h\big(c(t)~|~c(t-)\big)\big)}
\label{ercm}
\end{equation}
where $c(t-)\in \{0,1\}^{N \times (t-1)}$ is a matrix that indicates which cases have been citing in each other before time $t$, $\theta \in \mathbb{R}^q$ is a q-dimensional vector of parameters,  $h: \mathcal{C}_t(N) \to \mathbb{R}^q ~, ~ \c(t) \to (h_1(c(t)), \dots , h_q(c(t)))^T$ is a q-dimensional vector of different statistics and $\kappa(\theta) := \sum_{c(t)^*\in \mathcal{C}}\exp(\theta^T\cdot h(c(t)|c(t-)))$ is a normalization constant that ensures that (\ref{ercm}) defines a probability function on $\mathcal{C}_t$.\\[0.3cm]

@@

Our case for studying legal citaions at the directed dyadic level hinges upon the contribution to modeling offered by incorporating network dependence. To quantify this contribution, we use out-of-sample prediction [CITES]. Predicting out-of-sample offers an unbiased and general purpose way to evaluate the contribution, in terms of model fit, of one or more terms/parameters in a model. Unlike in-sample measures of model fit, out-of-sample methods are highly robust in avoiding overfitting, and work when we cannot accurately calculate the value of the likelihood function, as in the current case. Out-of-sample prediction is a common way to evaluate methods for modeling ties in networks, and has recently been applied to TERGMs in particular [CITE PA AND IEEE].
